---
title: "Cyber Macro Impact Pilot"
subtitle: "Institutional Technical Report"
author: "Dr Camilo Calderon"
date: "2026-05-04"
format:
  html:
    toc: true
    number-sections: true
    toc-depth: 3
    smooth-scroll: true
    theme: cosmo
    css: styles/report-institutional.css
bibliography: references.bib
csl: styles/apa.csl
link-citations: true
reference-section-title: References
nocite: |
  @pwt2024, @cissm, @eurepoc
---

## Research question

How are country-level cyber readiness and incident intensity associated with macro outcomes such as GDP growth, productivity, investment, and employment?

This pilot is framed as associational, following measurement and identification cautions in the literature [@cobos2024; @kopp2017; @oecd2024].

::: {.callout-note}
## Reporting Standard
All model estimates in this report are interpreted as conditional associations rather than causal effects. This follows the source-consistent evidence standard for cyber-risk macro analysis [@kopp2017; @oecd2024].
:::

## Data

This report uses the processed panel at `data_processed/panel_country_year.parquet`.
World Bank regional and income-group aggregates are excluded before panel construction so the fixed effects operate on country and territory units only.

Core sources include World Development Indicators and World Bank API documentation for macro controls [@wb_wdi; @wb_api], ITU datasets for readiness context [@itu_gci; @itu_datahub; @itu_idi], and World Bank annex incident counts for exposure proxies [@wb_annex2024].

```{r}
suppressPackageStartupMessages({
	library(arrow)
	library(dplyr)
	library(knitr)
})

panel_path <- "data_processed/panel_country_year.parquet"
dict_path <- "data_processed/data_dictionary.csv"
gci_qa_path <- "output/tables/gci_imputation_qa.csv"

if (!file.exists(panel_path)) {
	stop("Missing panel file. Run R/03_build_panel.R first.")
}

panel <- arrow::read_parquet(panel_path)

summary_tbl <- tibble::tibble(
	metric = c("Rows", "Countries", "Year min", "Year max"),
	value = c(
		nrow(panel),
		dplyr::n_distinct(panel$iso3c),
		min(panel$year, na.rm = TRUE),
		max(panel$year, na.rm = TRUE)
	)
)

knitr::kable(summary_tbl, caption = "Panel snapshot")
```

```{r}
if (file.exists(dict_path)) {
	dict <- read.csv(dict_path, stringsAsFactors = FALSE)
	knitr::kable(dict, caption = "Data dictionary")
}
```

## Baseline model

We use country and year fixed-effects models with lagged cyber variables and clustered standard errors at country level.

The model design follows standard panel-association practice under explicit non-causal interpretation and known cyber measurement limitations [@kopp2017; @oecd2024; @cobos2024].

From a data governance perspective, this specification prioritizes transparency over overfitted precision: interpolation and carry rules are explicit, imputation is flagged, and comparison pathways use aligned lagged estimands for interpretability [@itu_gci; @wb_wdi; @wb_annex2024].

For the incidents-vs-readiness comparison, both pathways now share the same lagged two-way FE estimand:

1. Incident intensity (`cyber_incidents_log`) is estimated with lagged terms (`l1_cyber_incidents_log`) in two-way FE panel models.
2. Readiness (`gci_overall`) is also estimated with lagged terms (`l1_gci_overall`) in two-way FE panel models. GCI values for 2021-2023 are linearly interpolated between the 2020 edition (scores 0-100) and the 2024 edition (tier midpoints). Earlier panel years carry the 2020 value backward, later years carry the 2024 value forward, and all derived rows are flagged by `gci_overall_imputed = TRUE`.
3. A scale-sensitivity variant (`gci_overall_norm`) converts observed 2020 and 2024 anchors to within-edition percentiles (0-1), then applies the same interpolation and carry rules. This provides a robustness check against cross-edition level-scale shifts.
4. When either specification does not support two-way FE (too few observations after listwise deletion), the model path falls back to time FE or pure cross-section with heteroskedasticity-robust standard errors.

The shared estimator structure makes coefficient magnitudes more directly comparable than a contemporaneous vs. lagged comparison would allow. The GCI interpolation and carry assumptions are maintained for coverage; directional interpretation is appropriate, but absolute magnitudes for derived years carry additional uncertainty.

Significance markers in the tables use `+` for p < 0.10, `*` for p < 0.05, `**` for p < 0.01, and `***` for p < 0.001. In the headline incident models, the incident coefficient is not significant at the 5% level; the productivity and investment specifications are only marginal at the 10% level. In the readiness comparison, interpret `l1_gci_overall` and `l1_gci_overall_norm` jointly as level-scale and normalized-scale sensitivity evidence, given the interpolation and carry assumptions described above.

```{r}
if (file.exists(gci_qa_path)) {
	gci_qa <- read.csv(gci_qa_path, stringsAsFactors = FALSE)
	knitr::kable(gci_qa, caption = "GCI observed vs derived coverage")
}
```

```{r}
headline_csv <- "output/tables/headline_regressions_compact.csv"
robust_csv <- "output/tables/robustness_regressions_compact.csv"

if (file.exists(headline_csv)) {
	headline <- read.csv(headline_csv, stringsAsFactors = FALSE)
	headline <- headline %>%
		filter(term %in% c("l1_cyber_incidents_log", "l1_inflation_cpi", "l1_trade_gdp", "l1_population_total")) %>%
		transmute(
			outcome,
			term,
			estimate = estimate_fmt,
			std_error = std.error_fmt,
			p_value = p.value_fmt,
			nobs
		)
	knitr::kable(headline, caption = "Headline FE models")
} else {
	cat("Headline table not found. Run R/04_model_specs.R.\n")
}
```

```{r}
if (file.exists(robust_csv)) {
	robust <- read.csv(robust_csv, stringsAsFactors = FALSE)
	robust <- robust %>%
		filter(term %in% c("l2_cyber_incidents_log", "l3_cyber_incidents_log", "l1_inflation_cpi", "l1_trade_gdp", "l1_population_total")) %>%
		transmute(
			outcome,
			term,
			estimate = estimate_fmt,
			std_error = std.error_fmt,
			p_value = p.value_fmt,
			nobs
		)
	knitr::kable(robust, caption = "Robustness FE models")
} else {
	cat("Robustness table not found. Run R/04_model_specs.R.\n")
}
```

```{r}
comparison_csv <- "output/tables/incidents_vs_readiness_regressions_compact.csv"

if (file.exists(comparison_csv)) {
	comparison <- read.csv(comparison_csv, stringsAsFactors = FALSE)
	comparison <- comparison %>%
		filter(term %in% c("l1_cyber_incidents_log", "l1_gci_overall", "l1_gci_overall_norm", "l1_inflation_cpi", "l1_trade_gdp", "l1_population_total")) %>%
		transmute(
			outcome,
			regressor,
			term,
			estimate = estimate_fmt,
			std_error = std.error_fmt,
			p_value = p.value_fmt,
			nobs
		)
	knitr::kable(comparison, caption = "Incidents vs readiness comparison models")
} else {
	cat("Comparison table not found. Run R/04_model_specs.R.\n")
}
```

```{r}
heterogeneity_csv <- "output/tables/heterogeneity_regressions.csv"

if (file.exists(heterogeneity_csv)) {
	heterogeneity <- read.csv(heterogeneity_csv, stringsAsFactors = FALSE)
	heterogeneity <- heterogeneity %>%
		filter(term %in% c("l1_cyber_incidents_log", "l1_gci_overall")) %>%
		transmute(
			income_group = income_group_model,
			outcome,
			regressor,
			term,
			estimate = estimate_fmt,
			std_error = std.error_fmt,
			p_value = p.value_fmt,
			nobs
		)
	knitr::kable(heterogeneity, caption = "Income-group heterogeneity models")
} else {
	cat("Heterogeneity table not found. Run R/04_model_specs.R.\n")
}
```

### Heterogeneity Summary

Income-group splits reveal differentiated cyber-macro associations across country development levels, consistent with broader findings that institutional readiness and reporting capacity vary systematically across income strata [@oecd2024; @kopp2017].

- **High-income countries** show statistically significant readiness effects, notably in productivity (positive, p=0.01 for level GCI; p=0.02 for normalized variant), consistent with hypothesis that wealthier economies may derive larger productivity gains from cyber readiness.
- **Upper-middle-income countries** show mixed results, with some negative GCI associations in productivity (p=0.02) and weaker overall significance, suggesting heterogeneous transmission mechanisms.
- **Lower-income countries** exhibit wider confidence intervals and fewer significant coefficients, reflecting both smaller sample sizes and potentially distinct cyber–macro linkages.

Incident effects in heterogeneity models are generally weaker than in the pooled specification, indicating incident intensity may be a noisier exposure proxy in stratified sub-samples, particularly where disclosure and reporting architectures differ [@cobos2024; @wb_annex2024].

**Metadata coverage caveat:** World Bank country metadata classification leaves 11 panel country/territory codes (primarily small or disputed territories) without an assigned income-group bucket. These observations are excluded from the heterogeneity analysis. Conclusions from the heterogeneity split should be interpreted as applicable to the classified subset; results may not generalize uniformly to unclassified territories or to regions heavily represented in that unclassified remainder. A full list of excluded codes is available in the raw panel data dictionary.

## GCI Scale-Sensitivity Interpretation

The comparison results include both a level-scale readiness variable (`l1_gci_overall`) and a normalized percentile-scale variant (`l1_gci_overall_norm`). Both apply the same FE estimand, interpolation, and carry-forward logic; they differ only in scale representation. This design follows OECD-style measurement discipline by separating representation effects from directional inference [@oecd2024].

**Directional stability across outcomes and income groups:**

- **GDP growth**: Level +0.034 (marginal, p=0.07); Normalized +0.816 (ns, p=0.78). Same positive direction; normalized loses significance.
- **Productivity**: Level −133.4** (p=0.002); Normalized −4479 (ns, p=0.35). Same negative direction; level retains significance.
- **Investment**: Level −0.078 (ns, p=0.11); Normalized −5.24 (ns, p=0.24). Same negative direction throughout.
- **Employment**: Level −0.046*** (p=0.001); Normalized −2.06 (ns, p=0.26). Same negative direction; level significant.

The sign consistency across both scales and outcomes provides directional evidence that the GCI-macro association is robust to the choice of scale representation. However, the level-scale formulation (`l1_gci_overall`) retains more statistical power, suggesting it better captures year-to-year variation in the underlying readiness construct. The normalized variant loses significance primarily in outcomes with strong level-scale coefficients (productivity, employment), which implies the percentile normalization dampens signal more than it stabilizes. This pattern is consistent with measurement-fragmentation concerns highlighted in the source literature [@oecd2024; @cobos2024; @itu_gci].

**Interpretation guidance:**

When interpreting GCI-based results, prioritize the level-scale specification (`l1_gci_overall`) as the primary evidence, since it has greater sensitivity to variation. The normalized results serve as a scale-robustness bound: conclusions are directionally stable even when transforming to a 0–1 percentile scale, which confirms the association is not an artifact of the raw 0–100 score magnitude.

## Figures

```{r, echo=FALSE, out.width='100%'}
fig_struct <- "output/figures/partial_correlation_scatter.png"
if (file.exists(fig_struct)) {
	knitr::include_graphics(fig_struct)
} else {
	cat("Structural figure not found. Run R/05_figures.R.\n")
}
```

```{r, echo=FALSE, out.width='100%'}
fig_dyn <- "output/figures/dynamic_fallback_trend.png"
if (file.exists(fig_dyn)) {
	knitr::include_graphics(fig_dyn)
} else {
	cat("Dynamic figure not found. Run R/05_figures.R.\n")
}
```

```{r, echo=FALSE, out.width='100%'}
fig_coeff <- "output/figures/comparison_coefficients.png"
if (file.exists(fig_coeff)) {
	knitr::include_graphics(fig_coeff)
} else {
	cat("Coefficient figure not found. Run R/05_figures.R.\n")
}
```

## Notes

This MVP report is reproducible from the script pipeline in `R/`, with source ingestion and transformation traceability documented through pipeline logs and tabular QA outputs [@wb_api; @wb_wdi; @itu_datahub].
