---
title: "Cyber Macro Impact Pilot"
format: html
---

## Research question

How are country-level cyber readiness and incident intensity associated with macro outcomes such as GDP growth, productivity, investment, and employment?

## Data

This report uses the processed panel at `data_processed/panel_country_year.parquet`.
World Bank regional and income-group aggregates are excluded before panel construction so the fixed effects operate on country and territory units only.

```{r}
suppressPackageStartupMessages({
	library(arrow)
	library(dplyr)
	library(knitr)
})

panel_path <- "data_processed/panel_country_year.parquet"
dict_path <- "data_processed/data_dictionary.csv"

if (!file.exists(panel_path)) {
	stop("Missing panel file. Run R/03_build_panel.R first.")
}

panel <- arrow::read_parquet(panel_path)

summary_tbl <- tibble::tibble(
	metric = c("Rows", "Countries", "Year min", "Year max"),
	value = c(
		nrow(panel),
		dplyr::n_distinct(panel$iso3c),
		min(panel$year, na.rm = TRUE),
		max(panel$year, na.rm = TRUE)
	)
)

knitr::kable(summary_tbl, caption = "Panel snapshot")
```

```{r}
if (file.exists(dict_path)) {
	dict <- read.csv(dict_path, stringsAsFactors = FALSE)
	knitr::kable(dict, caption = "Data dictionary")
}
```

## Baseline model

We use country and year fixed-effects models with lagged cyber variables and clustered standard errors at country level.

For the incidents-vs-readiness comparison, both pathways now share the same lagged two-way FE estimand:

1. Incident intensity (`cyber_incidents_log`) is estimated with lagged terms (`l1_cyber_incidents_log`) in two-way FE panel models.
2. Readiness (`gci_overall`) is also estimated with lagged terms (`l1_gci_overall`) in two-way FE panel models. GCI values for 2021-2023 are linearly interpolated between the 2020 edition (scores 0-100) and the 2024 edition (tier midpoints). Earlier panel years carry the 2020 value backward, later years carry the 2024 value forward, and all derived rows are flagged by `gci_overall_imputed = TRUE`.
3. When either specification does not support two-way FE (too few observations after listwise deletion), the model path falls back to time FE or pure cross-section with heteroskedasticity-robust standard errors.

The shared estimator structure makes coefficient magnitudes more directly comparable than a contemporaneous vs. lagged comparison would allow. The GCI interpolation and carry assumptions are maintained for coverage; directional interpretation is appropriate, but absolute magnitudes for derived years carry additional uncertainty.

Significance markers in the tables use `+` for p < 0.10, `*` for p < 0.05, `**` for p < 0.01, and `***` for p < 0.001. In the headline incident models, the incident coefficient is not significant at the 5% level; the productivity and investment specifications are only marginal at the 10% level. In the readiness comparison, `l1_gci_overall` is significant for productivity and employment, but these estimates rely heavily on the GCI interpolation and carry assumptions described above.

```{r}
headline_csv <- "output/tables/headline_regressions.csv"
robust_csv <- "output/tables/robustness_regressions.csv"

if (file.exists(headline_csv)) {
	headline <- read.csv(headline_csv, stringsAsFactors = FALSE)
	knitr::kable(headline, caption = "Headline FE models")
} else {
	cat("Headline table not found. Run R/04_model_specs.R.\n")
}
```

```{r}
if (file.exists(robust_csv)) {
	robust <- read.csv(robust_csv, stringsAsFactors = FALSE)
	knitr::kable(robust, caption = "Robustness FE models")
} else {
	cat("Robustness table not found. Run R/04_model_specs.R.\n")
}
```

```{r}
comparison_csv <- "output/tables/incidents_vs_readiness_regressions.csv"

if (file.exists(comparison_csv)) {
	comparison <- read.csv(comparison_csv, stringsAsFactors = FALSE)
	knitr::kable(comparison, caption = "Incidents vs readiness comparison models")
} else {
	cat("Comparison table not found. Run R/04_model_specs.R.\n")
}
```

## Figures

```{r, echo=FALSE, out.width='100%'}
fig_struct <- "output/figures/partial_correlation_scatter.png"
if (file.exists(fig_struct)) {
	knitr::include_graphics(fig_struct)
} else {
	cat("Structural figure not found. Run R/05_figures.R.\n")
}
```

```{r, echo=FALSE, out.width='100%'}
fig_dyn <- "output/figures/dynamic_fallback_trend.png"
if (file.exists(fig_dyn)) {
	knitr::include_graphics(fig_dyn)
} else {
	cat("Dynamic figure not found. Run R/05_figures.R.\n")
}
```

## Notes

This MVP report is reproducible from the script pipeline in `R/`.
